Emergent dual scaling of riverine biodiversity

### 2021-06-01 manuscript_pnas.Rmd
edits: correct paths

---
title: "Emergent dual scaling of riverine biodiversity"
output:
  pdf_document:
    latex_engine: xelatex
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: references.bib
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
---

@@

  # read gis files
  filename <- list.files(path = here::here('empirical/data_gis'), full.names = T)
  wsd_subset <- lapply(filename[str_detect(filename, "wsd_subset")], st_read)
  point_subset <- lapply(filename[str_detect(filename, "point_subset")], st_read)
  shape <- lapply(filename[str_detect(filename, "shape")], st_read)
  shape[[1]] <- st_set_crs(shape[[1]], st_crs(wsd_subset[[1]]))

  # read theory file
  dat_sim <- read_csv(here::here('theory/result/result_sim2020-11-10.csv'))

@@

  # for Bayes factor
  source(here::here('empirical/analysis_rlm_bf.R'))

  # for average predictive comparisons
  source(here::here('empirical/analysis_apcomp.R'))

  # Hokkaido data
  dat_hkd_site <- read_csv(here::here('empirical/data_out/data_hkd_site.csv'))

  # Midwest data
  dat_mw_site <- read_csv(here::here('empirical/data_out/data_mw_site.csv')) %>% 
    mutate(Year = format(as.Date(Date, format = '%m/%d/%Y'), "%Y"))

@@

We depicted branching ecosystems as a bifurcating network of connected habitat patches [@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014] (**Figure 1**). Habitat patches, or local communities, can be either non-branching or branching river sections with a unit length $l$, which defines the scale of local species interactions. Two parameters determine the geometric properties of simulated branching networks: the number of habitat patches $N_p$ (ecosystem size) and branching probability $P_b$ (ecosystem complexity). Each habitat patch is assigned to be a branching patch (including upstream terminals) with probability $P_b$ or a non-branching patch with probability $1-P_b$. In this framework, an individual branch is a consecutive series of non-branching patches terminated at a branching patch; therefore, the number of habitat patches in branch $q$, $n_{p,q}$, is a realization of a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$ [@teruiMetapopulationStabilityBranching2018] but conditional on $\sum_{q=1}^{N_b}n_{p,q} = N_p$ ($N_b$ is the number of branches in a network). The number of branches $N_b$ is identical to the number of branching habitat patches in a network, or the number of 'success' out of $N_p$ trials. This leads to an assumption that $N_b$ is a realization of a random variable drawn from a binomial distribution as $N_b \sim Binomial(N_p, P_b)$. This framework has two merits. First, it reflects observed patterns of branch length distribution, which is known to follow an exponential distribution (a continuous version of a geometric distribution) [@rodriguez-iturbeFractalRiverBasins2001]. Second, it preserves the fractal nature of branching patterns [@peckhamReformulationHortonLaws1999]. An alternative approach is the use of optimal channel networks [@carraroGenerationApplicationRiver2020a]. However, we employed our framework because branching probability $P_b$ is quantifiable in natural river networks, enabling us to compare theoretical predictions with empirical observations.

>>>>

---
title: "Emergent dual scaling of riverine biodiversity"
output: word_document
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: references.bib
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
---

@@

  # read gis files
  filename <- list.files(path = here::here('code_empirical/data_gis'), full.names = T)
  wsd_subset <- lapply(filename[str_detect(filename, "wsd_subset")], st_read)
  point_subset <- lapply(filename[str_detect(filename, "point_subset")], st_read)
  shape <- lapply(filename[str_detect(filename, "shape")], st_read)
  shape[[1]] <- st_set_crs(shape[[1]], st_crs(wsd_subset[[1]]))

  # read theory file
  dat_sim <- read_csv(here::here('code_theory/result/result_sim2020-11-10.csv'))

@@

  # for Bayes factor
  source(here::here('code_empirical/analysis_rlm_bf.R'))

  # for average predictive comparisons
  source(here::here('code_empirical/analysis_apcomp.R'))

  # Hokkaido data
  dat_hkd_site <- read_csv(here::here('code_empirical/data_out/data_hkd_site.csv'))

  # Midwest data
  dat_mw_site <- read_csv(here::here('code_empirical/data_out/data_mw_site.csv')) %>% 
    mutate(Year = format(as.Date(Date, format = '%m/%d/%Y'), "%Y"))

@@

We depicted branching ecosystems as a bifurcating network of connected habitat patches [@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014] (**Figure 1**). Habitat patches, or local communities, can be either non-branching or branching river sections with a unit length $l$, which defines the scale of local species interactions. Two parameters determine the geometric properties of simulated branching networks: the number of habitat patches $N_p$ (ecosystem size) and branching probability $P_b$ (ecosystem complexity). Each habitat patch is assigned to be a branching patch (including upstream terminals) with probability $P_b$ or a non-branching patch with probability $1-P_b$. In this framework, an individual branch is a consecutive series of non-branching patches terminated at a branching patch; therefore, the number of habitat patches in branch $q$, $n_{p,q}$, is a realization of a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$ [@teruiMetapopulationStabilityBranching2018] but conditional on $\sum_{q=1}^{N_b}n_{p,q} = N_p$ ($N_b$ is the number of branches in a network). The number of branches $N_b$ is identical to the number of branching habitat patches in a network, or the number of 'success' out of $N_p$ trials. This leads to an assumption that $N_b$ is a realization of a random variable drawn from a binomial distribution as $N_b \sim Binomial(N_p, P_b)$. This framework has two merits. First, it reflects observed patterns of branch length distribution, which is known to follow an exponential distribution (a continuous version of a geometric distribution) [@rodriguez-iturbeFractalRiverBasins2001]. Second, it preserves the fractal nature of branching patterns [@peckhamReformulationHortonLaws1999]. An alternative approach is the use of optimal channel networks [@carraroGenerationApplicationRiver2020]. However, we employed our framework because branching probability $P_b$ is quantifiable in natural river networks, enabling us to compare theoretical predictions with empirical observations.
